---
title: "Filters and joins"
author: "Heather Lowers"
date: "1/28/2020"
output: html_document
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE)

library(tidyverse)
library(janitor)
library(readxl)
library(here)
library(kableExtra)
```

Get some fish.csv and first worksheet (first sheet is default) of kelp_fronds.xlsx data.  Add sheet argument if the first worksheet is not the one you want to read in.  You can call it either by name or number.

```{r}
fish <- read_csv(here("data", "fish.csv"))
kelp_abur <- read_excel(here("data", "kelp_fronds.xlsx"), sheet = "abur")
```

### Filter by an exact match

Exact match: `==`

Example 1: only keep observations from `fish` where the common_name variable matches (==) "garibaldi"

```{r}
fish_garibaldi <- fish %>% 
  filter (common_name == "garibaldi")
```

>, <, >=, <=, ==  

Create a subset of fish called fish_over50, starting from fish, where the total_count is greater than OR equal to 50

```{r}
fish_50 <- fish %>% 
  filter (total_count >= 50)
```

### What if we want to filter by this OR this?

Use can use the vertical line `|` or "or"


```{r}
fish_2sp <- fish %>% 
  filter (common_name == "garibaldi"| common_name == "blacksmith")

# If you're filtering for multiple matches within the same variable, you can use %in%

fish_3sp <- fish %>% 
  filter (common_name %in% c("garibaldi","blacksmith","black surfperch"))

```

Activity: subset from fish (fish_gar_2016) that keeps all observations where the year is 2016 OR the common_name is "garibaldi"

```{r}
fish_gar_2016 <- fish %>% 
  filter (common_name == "garibaldi"| year == 2016)
```

### Filter to match this AND that
use "," or &

```{r}
fish_aque_2018 <- fish %>% 
  filter (year == 2018, site == "aque")
```

To exclude things: `!=`  ie. I want to include everything except 

### stringr::str_detect()  Used to filter based on partial matches

```{r}
fish_black <- fish %>% 
  filter(str_detect(common_name, pattern = "black"))
```

Subset of data if the common name must include the string "it"?

```{r}
fish_it <- fish %>% 
  filter(str_detect(common_name, pattern = "it"))
```

### `dplyr::*_join`

```{r}
ab_kelpfish <- kelp_abur %>% 
  full_join(fish, by = c("year", "site"))
```

### `left_join()` to keep all from x, and merge matching things from y with it (exclude things from y that don't have a match in x)

```{r}
kelp_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year", "site"))
```

### `inner_join()` to merge data frames, only keeping observations with a match in both

```{r}
kelp_fish_injoin <- kelp_abur %>% 
  inner_join(fish, by = c("year", "site"))
```

### Pipe together multiple steps

- create an object called `my_fish_join`
- starting from `fish`
- filtering to only including observations at site "abur" from the year 2017
- join the kelp_abur data frame to the resulting subset using left_join()
- add a new column (`mutate`) to calculate fish per kelp frond density (counts of fish/total fronds)

```{r}
my_fish_join <- fish %>% 
  filter(site == "abur", year == "2017") %>% 
  left_join(kelp_abur, by = c("year", "site")) %>% 
  mutate(fish_per_frond = total_count/total_fronds)
```

Nice HTML table using kable + kableExtra

```{r}
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```

Create a graph x is common_name and y is total 

```{r}
ggplot(data = my_fish_join, aes(x = common_name, y = fish_per_frond)) + 
  geom_col(aes(fill = common_name))
```

