---
title: "Creating graphs with ggplot2"
author: "Heather Lowers"
date: "1/27/2020"
output: html_document

---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.  

We will install these packages:

- tidyverse
- readxl: to read in excel files (in many ways)
- here: to help with file paths

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse) # install.pacakges("tidyverse")
library(readxl) #install.packages("readxl")
library(here) #install.packages("here")
```

# California National Parks Data
```{r read-data}

#read in and name it
ca_np <- read_csv(here("data/ca_np.csv"))
#also be written
#ca_np <- read_csv("data". "ca_np.csv")
#ca_np <- read_csv("data/ca_np.csv")

# also read in Channel Islands data
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```

### Make my own first graph!

We're going to make a line plot of year (x-variable) vs visitors (y-variable) in ggplot

Command + ALT + I is the keyboard shortcut to insert code chunk.

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_col()
# Update the geom_ to geom_point(), geom_col(), geom_area()
```

Making a base layer for my Channel Islands plots: 
```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

gg_base + geom_point()
```

### Now let's start customizing!

- line/point colors: color
- change line type: linetype
- change point style: pch
- change transparency: alpha

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

Activity: create a ggplot scatterplot (geom_point) that has triangle symbols (hint: ?pch), of a non-default size (size = #), and that are an updated color.
```{r}
gg_base +
  geom_point(color = "red",
             pch = 2,
             size = 4)
```


### Update graph aesthetics based on variables
```{r}
gg_base + 
  geom_point(
    aes(size = visitors, color = year), 
    pch = 2
  )

## Add arguments to update variables based on visitors and color
```

